---
title: Data Analytics Capstone Title
subtitle: "My Project"
author:
  - name: YOUR NAME 
    affiliation: SUNY Geneseo
    
  - name: YOUR COWORKER 1
    affiliation: SUNY Geneseo
    
date: last-modified

execute: 
  echo: false
  eval: true
  warning: false
  message: false
  fig-width: 10
  fig-height: 6

format:
  docx:
    toc: false
    number-sections: true
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)
library(rmarkdown)

# Add other packages as needed.
# Examples:
# library(broom)
# library(stargazer)
# library(modelsummary)
# library(glmnet)
# library(ranger)
# library(xgboost)

# Note: DT tables are designed mainly for HTML output.
# For Word output, use `kable()`, `gt`, `modelsummary`, or `stargazer`.
# Simple printed R output is also acceptable for basic summaries, but polished tables are preferred.
# It is also okay to add or revise tables manually in Microsoft Word after rendering the Word file from Quarto.

# Feel free to revise this ggplot theme_set
theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 90,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )

# Feel free to use below global options for color-blind-friendly scales
# scale_colour_discrete <- function(...) scale_colour_viridis_d(...)
# scale_fill_discrete <- function(...) scale_fill_viridis_d(...)
```

# Introduction

Describe the real-world motivation for your project. Clearly state your research question and explain why it matters.

Briefly preview your data, empirical approach, main findings, and contribution.

# Background and Related Literature

Summarize the relevant background, institutional setting, or prior research related to your topic.

Explain what is already known and how your project contributes to the existing discussion.

# Data

## Data Source

Describe the data source clearly. Include the source name, link, and access date in the References section.

## Unit of Observation

Explain what each row in your dataset represents.

## Key Variables

Describe your outcome variable, main explanatory variables, and important control variables.

## Data Cleaning

Explain major cleaning steps, sample restrictions, missing values, and variable construction.

## Summary Statistics

```{r summary-statistics}
# Replace this example with your own data and summary statistics.
# skim(your_data)
```

# Empirical Strategy / Methods

Describe the method used to answer your research question. This may include regression, classification, clustering, forecasting, text analysis, or another data analytics method.

For example, a **two-way fixed effects regression** can be used as an empirical strategy for **inference** when the goal is to estimate the relationship between a key explanatory variable and an outcome while accounting for unit-specific and time-specific factors:

$$
Y_{it} = \beta_0 + \beta_1 X_{it} + \delta' Z_{it} + \alpha_i + \lambda_t + \varepsilon_{it},
$$

where $Y_{it}$ is the outcome of interest for unit $i$ at time $t$, $X_{it}$ is the main explanatory variable, $Z_{it}$ represents control variables, $\alpha_i$ captures unit fixed effects, $\lambda_t$ captures time fixed effects, and $\varepsilon_{it}$ is the error term.

If your project uses machine learning or prediction, explain your train/test split, cross-validation, model tuning, and performance metrics.

# Results

Present your main empirical findings using tables, figures, and written interpretation.

```{r results-table}
# Replace this example with your own table or model output.
# modelsummary(list(model1, model2))
```

```{r results-figure}
# Replace this example with your own figure.
# ggplot(your_data, aes(x = x_variable, y = y_variable)) +
#   geom_point() +
#   labs(x = "X Variable", y = "Y Variable", title = "Main Result")
```

# Discussion and Implications

Interpret your results. Explain what they imply for economics, business, policy, or society.

Discuss limitations and clarify what your analysis can and cannot show.

# Conclusion

Summarize your research question, data, method, main findings, and key takeaway.

Briefly suggest possible future extensions.

# References

List all sources cited in your project, including data sources, academic articles, reports, websites, packages, and tools.

For online sources, include the full URL and date/time accessed.

# Appendix

Include additional tables, figures, robustness checks, code explanations, or variable definitions if needed.
